


```{r warning=FALSE message=False}
library('caTools')
library('ggplot2')
library('scales')

setwd('C:/Users/Propietario/Desktop/R/Coursera')
df <- read.csv('salaries.csv')
summary(df)
```

```{r}
#Splitting data set into training and testing sets
set.seed(356)
split <- sample.split(df$AnnualSalary, SplitRatio=3/4)
train <- subset(df, split == TRUE)
test <- subset(df, split == FALSE)

#Fitting simple linear regression to training and testing sets 
reg <- lm(formula = AnnualSalary ~ YearsOfExperience, data=train)
summary(reg)
```

```{r}
y_pred <- predict(reg, newdata=test)
summary(y_pred)
```

```{r}
#Visualizing Training set results
ggplot() + geom_point(aes(x=train$YearsOfExperience, y=train$AnnualSalary), color='red') +
      geom_line(aes(x=train$YearsOfExperience, y=predict(reg, newdata=train)), color='navy') +
      ggtitle('Anual Salaries of Data Scientist vs Years of Experience (Training set)') +
      xlab ('Years of Experience') +
      ylab ('Annual Salary') +
      scale_x_continuous(limits = c(0, 12)) +
      scale_y_continuous(limits = c(0, 150000))
```

```{r}
#trainig data points in red and in navy our prediction we just made, we can see we need to have 2.5 yoe to reach 50k

ggplot() + geom_point(aes(x=test$YearsOfExperience, y=test$AnnualSalary), color='red') +
      geom_line(aes(x=train$YearsOfExperience, y=predict(reg, newdata=train)), color='navy') +
      ggtitle('Anual Salaries of Data Scientist vs Years of Experience (Training set)') +
      xlab ('Years of Experience') +
      ylab ('Annual Salary') +
      scale_x_continuous(limits = c(0, 12)) +
      scale_y_continuous(limits = c(0, 150000))

#testing set visualized, lets say you are a data scientist eagerly looking for a job and applying to be hre, you know you have x yoe by implementing this regresion you can now predic your own salary if you are applying to a new job.
```

```{r}
test <- data.frame(YearsOfExperience = c(2, 4))
predic <- predict(reg, newdata= test)
predic
predic2 <- predict(reg, newdata= test, interval="confidence")
predic2
```

The output contains the following columns:

fit: the predicted sale values for the three new advertising budget
lwr and upr: the lower and the upper confidence limits for the expected values, respectively. By default the function produces the 95% confidence limits.

For example, the 95% confidence interval associated with a speed of 2 is (43698, 46837). This means that, according to our model, 2 years of experience will grant you a salary between 45 and 46.